30 years of privacy research for the masses \& \newline 15 years of decentralization attempts

### 2014-03-13 paper.tex
Label for Comparison section
Indent fixes
Removed the HEADS :P

	Although much research has been done about anonymous internet communication, only few systems have been actually implemented and used in practice. One of the most important factors that impact anonymity in a communication system is the number of users. A sufficiently large numbers of users are required for the system to make guarantees about its ability to conceal and hide users' communication. This makes it difficult to introduce a new anonymous internet system, due to the initial lack of users. Another important factor is for the system to be decentralized. Whenever a network has centralized components and a malicious third party gains control over these components, the functionality and perhaps even the anonymity of the system is compromised.

	The most widely used anonymous communication system is Tor. In this paper we will analyze Tor and its semi centralized nature. Furthermore we will look at attempts that have been made to decentralize Tor as well as several other decentralized systems.

	This paper is structured as following: In section \ref{sec:tor} we will give an introduction and overview about Tor. After that we will talk about decentralization and its problems in section \ref{sec:problems}. Vulnerabilities in Tor are discussed in section \ref{sec:attacks}. Afterwards, we will discuss the current state of decentralized internet systems in section \ref{sec:decentralized}. A comparison of existing decentralized networks is made in section \ref{sec:comparison}. Finally we will conclude and discuss our findings in section \ref{sec:conclusion}.

@@

	In 2002, it was decided to ditch the old code base and reimplement the project as Tor, the Second Generation Onion Router, introducing perfect forward secrecy, directory servers, hidden services and more \cite{dingledine2004tor}.

@@

		Say we have a circuit consisting of four nodes: our trusted client node ($C$), an (entry or guard) relay node ($X$), a (middle) relay node ($Y$) and an exit node ($Z$). In this case, there are three relay nodes, but this is not a fixed number. A visualisation of this path can be seen in figure \ref{fig:tor_layout}. Each node has its own public key and a corresponding private key. When building the circuit, our client generates a distinct secret for each of these nodes.

@@

	\subsection{Relay and exit nodes}
		The Tor network consists of several components. The clients in the Tor network are known as Onion Proxies. The software to run an Onion Proxy is available for free on the Tor website \cite{torprojectwebsite} and is easy to configure. The Onion Proxies are responsible for downloading the directory information, establish circuits across the network and handle connections from user applications.

		The routing in the network is done by Onion Routers, also called relay nodes. The relay nodes relays the data from the Onion Proxy to the web server across a circuit (circuits are described in 2D). Each Onion Router is connected to every other Onion Router with a TLS connection \cite{tlsprotocol}. Each circuit has three type of Onion Routers \cite{mccoy2008shining}:

@@

		As described in the previous section, data on the Tor networks travels over circuits. The data travels over the circuit in fixed-size cells that are 512 bytes long. Each cell has a header and a payload. The header consists of a number that identifies the circuit of the cell and a command that indicates what to do with the cell’s payload.

@@

		Suppose Alice is an Onion Proxy that wants to connect through the Tor network to a web server. Circuit setup uses the Diffie-Hellman key-exchange protocol \cite{diffiehellman} to establish a shared secret between nodes. To create a new circuit, Alice first sends a \emph{create} cell with the first half of the Diffie-Hellman handshake ($ g^b $) to the first node in her selected path (for example, OR1). OR1 sends a \emph{created} cell back with the second half of the key ($ g^b $) along with a hash of the final key. Now both Alice and OR1 have a shared key they use to encrypt and decrypt data between them.

@@

		First of all, the congestion control does not work well. The network has some problems handling bulk transfers, such as when downloading large files or streaming high-quality videos. The congestion control could be improved by using an unreliable protocol for links between Tor relays.

		Some Tor users put more traffic on the network than they contribute by running a Onion Router. This means that these users are slowing the network down as they use more traffic than giving to the network. A possible solution for this is to throttle certain protocols at exit nodes or at Onion Proxies.

		Also, the Tor network doesn't have the capacity to handle all the users that want privacy on the Internet. By increasing the amount of Onion Routing in the Tor network, the capacity is increased. Incentives such as LIRA \cite{jansen13lira} could make more users run a Onion Router, thus increasing the capacity of the network.

		The current path selection algorithm of Tor doesn't distribute the load evenly over the network. The problem is that the current selection strategy is optimal when the network is fully loaded. Using a better path selection algorithm could increase the capacity of the network and the overall user experience.

@@

		Much of the overhead of the network is in downloading the directory information. There is also overhead in the TLS connection between the nodes in the network. Removing the empty TLS application record which could reduce the overhead in the TCP/IP header with 6.3\%.

@@

		\subsection{Bootstrapping new nodes}
			If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be one starting point where this new node can discover neighbors in the network to connect with. In Tor a directory server can tell the new node what his neighbors are and how to find them \cite{dingledine2004tor}.

@@

			There are three types of file keys in Freenet: the first one is called the Keyword-Signed Key (KSK) which is derived from a short description of the file. Another key is the Signed-Subspace key (SSK) which enables personal namespaces. This key contains a public key and a private key where the private key is used to store the data and the public key is used in the queries of the file. The third type of key is the Content-Hash Key (CHK) which is used for updating and splitting of contents.

@@

			MIX networks typically are high latency and traffic moves slow. Messages will have to be stored on a node until enough messages have arrived to start sending them out in a random order. Often times cover traffic is used to generate enough messages for the network to obfuscate the real communication, which in turn generates a lot of bandwidth overhead. MorphMix has been developed as a low latency and high performance network. As such, it will not hold messages for very long, nor will it use cover traffic.

@@

@@

	\section{Comparison existing decentralized networks}
		In the last section, we discussed existing decentralized peer-to-peer networks. In this section we will make a comparison between these networks. For each network, we will explain the following features:

@@

@@

		In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore we looked at the main disadvantages and problems Tor currently has. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.

>>>>

	Although much research has been done on anonymous internet communication, only few systems have been actually implemented and are being used in practice. One of the most important factors that impact anonymity in a communication system is the number of users. A sufficiently large numbers of users are required for the system to make guarantees about its ability to conceal and hide users' communication. This makes it difficult to introduce a new anonymous internet system, due to the initial lack of users. Another important factor that impacts the anonymity of the system is decentralization. Whenever a network has centralized components and a malicious third party gains control over these components, the functionality and perhaps even the anonymity of the system is compromised.

	The most widely used anonymous communication system is Tor. In this paper we will analyze Tor and its semi-centralized nature. Furthermore we will look at attempts that have been made to decentralize Tor as well as several other decentralized systems.

	This paper is structured as following: In section \ref{sec:tor} we will give an introduction and overview of Tor. After that, we will talk about decentralization and its problems in section \ref{sec:problems}. Vulnerabilities in Tor are discussed in section \ref{sec:attacks}. Afterwards, we will discuss the current state of decentralized internet systems in section \ref{sec:decentralized}. A comparison of existing decentralized networks is made in section \ref{sec:comparison}. Finally, we will conclude and discuss our findings in section \ref{sec:conclusion}.

@@

	In 2002, it was decided to ditch the old code base and reimplement the project as Tor, the Second Generation Onion Router, introducing perfect forward secrecy, directory servers, hidden services and more \cite{dingledine2004tor}. In this section, we will explain the various compoments of Tor such as Onion Routing, the structure of the network, circuit creation and disadvantages of Tor.

@@

		Say we have a circuit consisting of four nodes: our trusted client node ($C$), an (entry or guard) relay node ($X$), a (middle) relay node ($Y$) and an exit node ($Z$). In this case, there are three relay nodes, but this is not a fixed number. A visualisation of this path can be seen in figure \ref{fig:tor_layout}. Each node has its own public key and a corresponding private key. When building the circuit, our client generates a distinct secret for each of these nodes. More information about the circuit setup, can be found in II-D. Circuit creation.

@@

	\subsection{Relay and exit nodes}
		The Tor network consists of several components. The clients in the Tor network are known as Onion Proxies. The software to run an Onion Proxy is available for free on the Tor website \cite{torprojectwebsite} and is easy for users to configure. The Onion Proxies are responsible for downloading the directory information, establishing circuits across the network and handling connections from user applications.

		The routing in the network is done by Onion Routers, also called relay nodes. The relay nodes relay the data from the Onion Proxy to the web server across a circuit (circuits are described in II-D). Each Onion Router is connected to every other Onion Router with a TLS connection \cite{tlsprotocol}. Each circuit has three type of Onion Routers \cite{mccoy2008shining}:

@@

		As described in the previous section, data on the Tor networks travels over circuits. The data travels over the circuit in fixed-size cells that are 512 bytes long. Each cell has a header and a payload. The header consists of a number that identifies the circuit the cell travels on and a command that indicates what to do with the cell’s payload.

@@

		Suppose Alice is an Onion Proxy that wants to connect through the Tor network to a web server. Circuit setup uses the Diffie-Hellman key-exchange protocol \cite{diffiehellman} to establish a shared secret between nodes. To create a new circuit, Alice first sends a \emph{create} cell with the first half of the Diffie-Hellman handshake ($ g^b $) to the first node in her selected path (for example, OR1). OR1 sends a \emph{created} cell back with the second half of the key ($ g^b $) along with a hash of the final key. Now both Alice and OR1 have a shared key they use to encrypt and decrypt data sent between them.

@@

		First of all, the congestion control does not work well. The network has some problems handling bulk transfers, such as when downloading large files or streaming high-quality videos. The congestion control could be improved by using an unreliable protocol for links between Tor relays. Goldberg et al. [linkje] have proposed PCTCP which could improve the response time of the Tor network with 60\% and the download time of files with 30\%.

		Some Tor users put more traffic on the network than they contribute by running a Onion Router. This means that these users are slowing the network down as they use more traffic than giving to the network. A possible solution for this is to throttle certain high-bandwidth protocols such as BitTorrent at exit nodes or at Onion Proxies.

		Also, the Tor network doesn't have the capacity to handle all the users that want privacy on the Internet. By increasing the amount of Onion Routing in the Tor network, the capacity is increased. Incentives such as LIRA \cite{jansen13lira} could make more users run a Onion Router, thus increasing the capacity of the network and making it faster.

		The current path selection algorithm of Tor doesn't distribute the load evenly over the network. The problem is that the current selection strategy is optimal when the network is fully loaded. This is not always the case. Using a better path selection algorithm could increase the capacity of the network and the overall user experience.

@@

		Much of the overhead of the network is in downloading the directory information. There is also overhead in the TLS connection between the nodes in the network. According to Dingledine et al, removing the empty TLS application record could reduce the overhead in the TCP/IP header with 6.3\%.

@@

		As we saw in the previous section, we would like to have a decentralized anonymous network. This isn't an easy task: 15 years of decentralization attempts on the Tor network are behind us. Several well-used decentralized peer-to-peer overlay networks are active such as Gnutella [link], Freenet [link] and Tapestry [link]. These networks will be described in more detail in the next chapter. Among these networks are some overlay anonymous networks such as Tarzan and Torks [link].

		\subsection{Bootstrapping new nodes}
			If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be a starting point where this new node can discover neighbors in the network to connect with. In Tor, a directory server can tell the new node what his neighbors are and where to find them \cite{dingledine2004tor}.

			Moving this system to a non peer-to-peer base is difficult: Dingledine et al. stated that this indeed is still an open problem. With decentralized systems, there is no central directory server to tell a new node where to locate neighbors and there might be a lot of malicious nodes in the network. Some systems such as Tarzan, MorphMix and Pastry \cite{rowstron2001pastry, rennhard2002introducing} are decentralized but they do suffer from performance issues. Some of these issues can be found in the last section when comparing these systems.

		\subsection{Incentives in decentralized systems}
			As described in section \ref{ss:tor_disadvantages}, people are usually using more bandwidth than that they donate. There are multiple proposals \cite{dingledine2010building, jansen13lira} to introduce incentives into Tor. If one would also have to build an incentive system into a decentralized system, they would have to find a way to manage the ratings of each client in the network, in such a way that they cannot be falsely modified. In other words, the reputation data has to be accurate and reliable. Besides the integrity of this data, the traffic it generates on the network should have minimal impact on the overall performance.

			Rahman \cite{rahman2009survey} proposes several options to build in incentives in a peer-to-peer network. The first proposal that is described is the so called \emph{Warm-glow Model}. This model determines the percentage of free-rides based on the probabilistic population distribution. If the percentage is above a certain threshold, the system will show signs of diminishing marginal returns.

			The third proposal is \emph{Reciprocity-Based Schemes}. Using this approach, a peer maintains a behavior history of other peers in the network. These schemes can be based on two somewhat reciprocities: direct reciprocity or indirect reciprocity. The formal are more suitable for longer relationships between peers. The latter is more scalable but they rely on third party and must handle trust issues themselves.

		\subsection{Key exchange}
			With a decentralized network, using a centralized authority for managing the keys, is not possible. This means that for secure communication, peers have to exchange the keys directly with each other, without a trusted party between them. Diffie-Hellman is a very popular algorithm for exchanging keys between two parties. Diffie-Hellman is for example used in the circuit setup in Onion Routing (see secion II-D), however, it is possible for an adversary to manipulate with the keys send between two parties, making the protocol vulnerable for a man-in-the-middle attack. This weakness makes it possible for an adversary to decrypt all message sent between the two parties.

			Tor currently uses an interactive forward-secret key-exchange protocol called the \emph{Tor Authentication Protocol} (TAP) [linkje naar ace]. This protocol uses telescoping which means that the initiator negotiates session keys with each successive hop in the circuit. There are several proposals for more efficient key exchange methods. One of them is \emph{ACE}, an one-way authenticated key exchange protocol. The authors of this methods claim to have an 46\% efficiency improvement on the side of the client and an efficiency improvement of nearly 19\% on the side of the Onion Routers. ACE requires clients to send one extra element in the key exchange. This however, does not introduce any overhead because the element fits in the unused space in a cell.
=======
	\subsection{Bootstrapping new nodes}
		If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be one starting point where this new node can discover neighbors in the network to connect with. In Tor a directory server can tell the new node what his neighbors are and how to find them \cite{dingledine2004tor}.

@@

		Many other decentralized peer-to-peer networks have been proposed and some have been implemented [linkjes naar papers]. Examples of fully decentralized networks are Gnutella, Freenet, Pastry, Tapestry, and Tarzan. Each of these decentralized networks will be discussed in the next section. Note that we're only considering fully decentralized networks here. Networks that are partially or hybrid centralized, such as KaZaA or Napster, are not discussed here.[link naar survey p2p?]

		\subsection{Gnutella}
			Gnutella is a decentralized peer-to-peer network used for distributed search of files. Since the network is fully decentralized, peers in the network are called 'servents', a combination of the words servers and clients. Each peer can act both as a server, answering queries, or as a client, requesting and executing search queries. In order for a client to bootstrap, a new peer connects to one of several known hosts that are almost always available. These peers can for example be found on http://gnutellahosts.com. Once the peer has joined, there are several messages a servant can send out:

			\begin{itemize}
				\item A peer sends a PING message to it's neighbours to announce it's presents. This message is forwarded to other peers and each peer sends a PONG message back. This way, the newly joined peer can learn about his neighbours and the network topology.
				\item The peer can issue QUERY answers. Other peers responds with a QUERY RESPONSE message to specify whether the file that was issued in the query, was found or not.
				\item To transfer items between peers, the GET and PUSH messages are being used.
			\end{itemize}

			Gnutella is an unstructured network which means that the placement of data items is not based on any knowledge of the topology nor the contents of the file. To search for a file, a flooding algorithm is used: each peer sends the request from another peer to each of it's neighbours in his routing table.
=======
		There are three types of file keys in Freenet: the first one is called the Keyword-Signed Key (KSK) which is derived from a short description of the file. Another key is the Signed-Subspace key (SSK) which enables personal namespaces. This key contains a public key and a private key where the private key is used to store the data and the public key is used in the queries of the file. The third type of key is the Content-Hash Key (CHK) which is used for updating and splitting of contents.

@@

			There are three types of file keys in Freenet: the first one is called the \emph{Keyword-Signed Key} (KSK) which is derived from a short description of the file. Another key is the \emph{Signed-Subspace key} (SSK) which enables personal namespaces. This key contains a public key and a private key where the private key is used to store the data and the public key is used in the queries of the file. The third type of key is the \emph{Content-Hash Key} (CHK) which is used for updating and splitting of contents.

			The routing algorithm for storing and retrieving data is dynamically and can adjust to the topology of the network. Each peer has only knowledge about his direct neighbours. Each request has a Hops-To-Live timer which indicates how many peers the request may traversal. Each peer decrements the timer by one and when the timer reaches zero, the request isn't forwarded any more. Results of queries are being cached in intermediate nodes to reduce the time for a query response. To prevent looping of the requests, each request contains a random identifier. The peers the request travels through, keep track of these identifiers and rejects the request if the request has already been answered by the peer.

		\subsection{MorphMix}
			MorphMix \cite{rennhard2002introducing} functions similar to both Tor and MIX networks. It relies on nested encryption and routing traffic over multiple nodes to ensure anonymity of its users' communication. Additionally, MorphMix uses the typical behavior of a MIX network, where it reorders messages that enter a node before sending them out.

			MIX networks are typically high latency networks and traffic in it moves slow. Messages will have to be stored on a node until enough messages have arrived to start sending them out in a random order. Often times cover traffic is used to generate enough messages for the network to obfuscate the real communication, which in turn generates a lot of bandwidth overhead. MorphMix has been developed as a low latency and high performance network. As such, it will not hold messages for very long, nor will it use cover traffic.

			In contrast to Tor, MorphMix does not feature a centralized node discovery mechanism. Instead, every node is free to chose a set of next nodes that will be the continuation of an anonymous tunnel. A malicious node could select a colluding node to continue the tunnel and therefore control the entire tunnel. To prevent this, a witness node is appointed to mediate the setup of an anonymous tunnel. Although this makes the possibility of such an attack more difficult, it doesn't make it impossible.

			An important part of MorphMix is the ability to detect malicious tunnels. Tunnels that are set up by well behaving nodes will select the next nodes in the tunnel randomly. Malicious colluding nodes, however, will specifically select nodes that are part of the malicious network to continue the tunnel. This will reveal itself in the fact that the probability of the selection of certain nodes is increased. By using this information it becomes possible, to an extent, to detect malicious tunnels.
=======
		MIX networks typically are high latency and traffic moves slow. Messages will have to be stored on a node until enough messages have arrived to start sending them out in a random order. Often times cover traffic is used to generate enough messages for the network to obfuscate the real communication, which in turn generates a lot of bandwidth overhead. MorphMix has been developed as a low latency and high performance network. As such, it will not hold messages for very long, nor will it use cover traffic.

@@

			Tapestry is a fundamental component of OceanStore [linkje naar paper over OceanStore], a decentralized storage system. Tapestry is also used in systems such as Bayeux and SpamWatch, a decentralized spam-filtering system.
=======
	\subsection{Tapestry}
		Tapestry is based on the Plaxton mesh data structure [linkje naar plaxton paper], which maintains pointers to node in the network whose IDs match the elements of a tree-like structure of ID prefixes up to a digit position. A property of Tapestry is that it offers load distribution and routing locality.

@@

\section{Comparison existing decentralized networks}
	\label{sec:comparison}

	In the last section, we discussed existing decentralized peer-to-peer networks. In this section we will make a comparison between these networks. For each network, we will explain the following features:

@@

			By running a HTTP exit node and a Tor router that eventually will act as an entry node in the network, an adversary can discover the identity of a user. The exit node injects an invisible iframe containing some JavaScript into the web pages that pass by, each sending a unique ID to a malicious web server. Every ten minutes, the Tor client chooses a new circuit and eventually an unlucky Tor client picks and uses the malicious entry node that was placed in the network.
=======
	\subsection{Sybil attacks}
		Hier wat over sybil attacks

			The solution for the JavaScript injection attack is disabling active content systems in the browser. For the HTML only variant, one would have to use HTTPS to prevent man-in-the-middle attacks.

@@

		In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore, we looked at the main disadvantages and problems Tor is currently facing. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.

		From this table, we come to the conclusion that there is no complete decentralized system that provides the full functionality and security that Tor offers today. While there is a lot of research on decentralized systems, there is not a system yet that can compete with Tor. The systems that do exist such as Tarzan, Torsk and Gnutella are getting things right, yet they often lack functionality such as anonymity, good routing performance or are vulnerable for a variety of attacks.

	\appendices
=======
	In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore we looked at the main disadvantages and problems Tor currently has. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.
